Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2021-09-22 methods_paper/fabl-paper.tex
add citations.

Record linkage is the task of identifying duplicate records across multiple data sources, often in the absence of a unique identifier \citep{christen2012data}. Record linkage is an increasingly important task in ``data cleaning," that is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others. Many statistical record linkage methods are the results of extensions from the groundbreaking work of \cite{fs, newcombe}. Specifically, \cite{newcombe} created comparison vectors for each pair of records in the data and independently classified those pairs as a match or a non-match. While this independence assumption may not always be reasonable, it is appealing for its mathematical simplicity and relative computational ease. Many recent work in the statistical literature has extended the aforementioned work [Brian: add citations here: Winkler and Thibadeau, 1991, Fair, 2004, Wagner and Layne, 2014, Gill and Goldacre, 2003, Singleton, 2013).

In the literature, bipartite record linkage is commonly known as merging two databases that contain duplication across the two databases, but do not contain any duplications within each database \citep{sadinle2017}.

@@

Despite the appeal of independent classification, it often leads to linkage decisions that do not enjoy transitive closure, and ignores that fact that we are often in scenarios where we know one record in one file has at most one match in another file. This setting is known as bipartite matching, and has received much less attention in the literature. \citep{sadinle2017} developed a prior distribution on the space of bipartite matchings such that his method, known as Beta Record Linkage (BRL), resulted in one-to-one matchings without any need for post-processing. However, BRL was even more computationally intensive than the original Fellegi-Sunter method, making it infeasible for large linkage tasks.

In this paper, we propose an extension to the \texttt{BRL} method proposed by \citep{sadinle2017} for this very end.  We increase the speed of the \texttt{BRL} procedure through a modification to the model specification that allows for parallel computing of the linkage parameter, and use hashing techniques to hasten calculations and reduce computational complexity. Additionally, we introduce storage efficient indexing (SEI), and a data partitioning approach to make the method amenable to large linkage tasks. To recognize the lineage from the original \texttt{BRL} method, we name our method \emph{Fast Beta Linkage} (\texttt{fabl},
pronounced ``fable'')

@@

\citep{sadinle2017} uses the same comparison vector approach as Fellegi and Sunter, but proposed a prior distribution for the set of matches that strictly enforced one-to-one matching without the need for any post-hoc processing procedures. Specifically, in each iteration of his Gibbs sampler, he considers each record \(j\in B\), removes from consideration the records \(i\in A\) that have already been matched, and then samples a potential link. This leads to a sampler that is significantly more accurate than the standard Fellegi-Sunter approach. However,
accounting for these dependencies throughout the linkage process is
computationally burdensome, leaving \texttt{BRL} only suitable for small
to moderate linkage problems.

@@

In order to make fair comparisons against the \citep{sadinle2017} model, we
integrate over the posterior of \(\lambda\) and rearrange terms to
produce the final full conditional:

@@

For a more in-depth explanation of this function and the induced Bayes
estimate, see \citep{sadinle2017}.

@@

\bibliographystyle{jasa}
{
\footnotesize
\bibliography{references}
}

>>>>

Record linkage is the task of identifying duplicate records across multiple data sources, often in the absence of a unique identifier \citep{christen_2012}. Record linkage is an increasingly important task in ``data cleaning," that is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others. Many statistical record linkage methods and case studies are extensions of seminal work of \cite{fellegi_theory_1969} and \cite{newcombe_automatic_1959}. Specifically, \cite{fellegi_theory_1969} created comparison vectors for each pair of records in the data and independently classified those pairs as a match or a non-match using a likelihood ratio test (or hypothesis test). Recent work in the statistical literature has extended the aforementioned work \citep{winkler1991application, fair2004generalized, wagner2014person, gill2003english}.

%While this independence assumption may not always be reasonable, it is appealing for its mathematical simplicity and relative computational ease. Many recent work in the statistical literature has extended the aforementioned work [Brian: add citations here:

In this paper, we consider bipartite record linkage, which is commonly known as merging two databases that contain duplication across the two databases, but do not contain any duplications within each database \citep{sadinle_bayesian_2017}. Much of the statistical literature focuses on bipartite record linkage \citep{fellegi_theory_1969, Jaro1989, Winkler1988, belin_1995, larsen_2001, liseo_2011,  herzog2007data, gutman_bayesian_2013, sadinle_bayesian_2017}.

@@

Despite the appeal of independent classification, it often leads to linkage decisions that do not enjoy transitive closure, and ignores that fact that we are often in scenarios where we know one record in one file has at most one match in another file. This setting is known as bipartite matching, and has received much less attention in the literature. \citep{sadinle_bayesian_2017} developed a prior distribution on the space of bipartite matchings such that his method, known as Beta Record Linkage (BRL), resulted in one-to-one matchings without any need for post-processing. However, BRL was even more computationally intensive than the original Fellegi-Sunter method, making it infeasible for large linkage tasks.

In this paper, we propose an extension to the \texttt{BRL} method proposed by \citep{sadinle_bayesian_2017} for this very end.  We increase the speed of the \texttt{BRL} procedure through a modification to the model specification that allows for parallel computing of the linkage parameter, and use hashing techniques to hasten calculations and reduce computational complexity. Additionally, we introduce storage efficient indexing (SEI), and a data partitioning approach to make the method amenable to large linkage tasks. To recognize the lineage from the original \texttt{BRL} method, we name our method \emph{Fast Beta Linkage} (\texttt{fabl},
pronounced ``fable'')

@@

\citep{sadinle_bayesian_2017} uses the same comparison vector approach as Fellegi and Sunter, but proposed a prior distribution for the set of matches that strictly enforced one-to-one matching without the need for any post-hoc processing procedures. Specifically, in each iteration of his Gibbs sampler, he considers each record \(j\in B\), removes from consideration the records \(i\in A\) that have already been matched, and then samples a potential link. This leads to a sampler that is significantly more accurate than the standard Fellegi-Sunter approach. However,
accounting for these dependencies throughout the linkage process is
computationally burdensome, leaving \texttt{BRL} only suitable for small
to moderate linkage problems.

@@

In order to make fair comparisons against the \citep{sadinle_bayesian_2017} model, we
integrate over the posterior of \(\lambda\) and rearrange terms to
produce the final full conditional:

@@

For a more in-depth explanation of this function and the induced Bayes
estimate, see \citep{sadinle_bayesian_2017}.

@@

\bibliographystyle{jasa}
{
\footnotesize
\bibliography{biblio}
}
